---
title: "Unit 9 HW"
author: 'MSDS 6372: Jacob Turner'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Conceptual Questions
A.  State the assumptions of the LDA model.
B.  What is the fundamental difference between the LDA and QDA model?
C.  What is a confusion matrix? 
D.  What is the fundamental difference between sensitivity and precision (positive predictive value)?

## Exercise 2: Decision Boundary Investigation
Consider the following data set:  
```{r cars}
library(RSSL)
set.seed(1234)
mydata<-generateCrescentMoon(200,2,2)
mydata$Class<-factor(ifelse(mydata$Class=="+","Yes","No"))
plot(mydata$X1,mydata$X2,col=mydata$Class,asp=1,pch=16,cex=.7,xlab="X1",ylab="X2")

```

A.  Would a KNN model or LDA/QDA model fare better for this example? State your reasons.  If you are stumped, complete part B and come back.

B.  The following code fits a KNN model using the `caret` package and provides a visualization of the prediction boundary.  Repurpose the code to see what the decision boundary is for an LDA model and a QDA model.
```{r}
library(caret)
fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
set.seed(1234)

#knn
knn.fit<-train(Class~X1+X2,
               data=mydata,
               method="knn",
               trControl=fitControl,
               metric="logLoss")


#Creating a validation set across the whole graph
validate <- expand.grid(X1=seq(-10, 10, by=0.25), # sample points in X1
                  X2=seq(-15, 15, by=0.5)) # sample points in X2

#Computing predicted probabilities on the training data
predictions <- predict(knn.fit, validate, type = "prob")[,"Yes"]

threshold=0.5
class_pred = as.factor(ifelse(predictions>threshold, "1", "0"))

color_array <- c("black", "red")[as.numeric(class_pred)] 

par(mfrow=c(1,2))
plot(mydata$X1,mydata$X2,col=mydata$Class,asp=1,pch=16,cex=.7,xlab="X1",ylab="X2")
plot(validate,col=color_array,pch=20,cex=0.15)
par(mfrow=c(1,2))




```

## Exercise 3: LDA and thresholding practice
The cancer data set contains multiple genetic variables in their bloood measured on patients with tumors in the throat/esophagus. After surgery, the tumors were biopsied and classified as malignant or benign.  The data set actually has 4 data sets merged together from different hospitals. For this exercise, the code below is going to collapse them into a sing train/validation split. The censor variable is the response code as 0 and 1.  

```{r pressure, echo=FALSE}
setwd("~/Desktop/CourseDev/Unit 9/HW")
cancer<-read.csv("CancerExample.csv",stringsAsFactors = T)
index<-which(cancer$Set=="Training")
index2<-which(cancer$Set=="Validation I")
training<-cancer[index,c(5:14)]
validate<-cancer[index2,c(5:14)]
```

A.  Recode the response variable so that it is a factor with levels Benign (cancer) and Malignant (healthy). Make sure the recoded response exists in both training and validation sets. Provide a `ggpairs` plot of the nine genetic predictor variables by cancer status on the training data. Do the assumptions of LDA look reasonable?  Does it look like the predictors will be able to predict the response effectively given how LDA tries to classify? Provide commentary and offer a suggestion to fix issues if you mention any. 

B.  Compute an LDA model on the training data and compute the confusion matrix on the validation set using a threshold of 0.5.  Underneath this setting is the classification more specific or more sensitive? These data sets are not random samples so we can not really look into PPV and NPV that closely.  Compute a second confusion matrix by changing the threshold to get something that has an even balance between sensitivity and specificity. Note: This is not about reading into the values but rather working through the computational steps.

## Exercise 4:  Curse of Dimensionality
The following code simulates a training and validation set from multivariate normal distributions.  The first two predictors are important  while the remaining predictors are just random noise. I've included a plot of the first five for verification.
```{r}
library(mvtnorm)
set.seed(1234)
muYes<-c(10,10)
muNo<-c(8,8)
Sigma<-matrix(c(1,.8,.8,1),2,2,byrow=T)
nY<-30
nN<-30
dataYes<-rmvnorm(nY,muYes,Sigma)
dataNo<- rmvnorm(nN,muNo,Sigma)
train<-rbind(dataYes,dataNo)
train<-data.frame(train)
for (i in 3:20){
 train<-cbind(train,rnorm(nY+nN))
}
names(train)<-paste("X",1:20,sep="")

train$Response<-rep(c("Yes","No"),each=30)
train$Response<-factor(train$Response)

#Creating a validation set
muYes<-c(10,10)
muNo<-c(8,8)
Sigma<-matrix(c(1,.8,.8,1),2,2,byrow=T)
nY<-500
nN<-500
dataYes<-rmvnorm(nY,muYes,Sigma)
dataNo<- rmvnorm(nN,muNo,Sigma)
validate<-rbind(dataYes,dataNo)
validate<-data.frame(validate)
for (i in 3:20){
  validate<-cbind(validate,rnorm(nY+nN))
}
names(validate)<-paste("X",1:20,sep="")

validate$Response<-rep(c("Yes","No"),each=500)
validate$Response<-factor(validate$Response)

pairs(train[,1:5],col=rep(c("red","black"),each=30))
```

A.  Fit an LDA model on the training data using just the first two predictions $X1$ and $X2$.  Observe the overall accuracy on the validation set.  (Do not worry about the other metrics for this exercise, but it might be interesting to examine those as well.)

B.  Fit an additional LDA model using the first 10 predictors which will include eight random noise predictors. Observe the overall accuracy on the validation set.

C.  Repeat again one last time using all 20 predictors (18 random noise predictors) and observe the overall accuracy on the validation set.  

D.  What trend do you see in the accuracy as you add more predictors that are just random noise?  This is referred to as the "curse of dimensionality" and also occurs in other algorithms such as KNN.  It highlights the importance of feature selection in the case of LDA. To make matters worse, these tools do not traditionally have feature selection procedures built in so we have to consider other tools for feature selection and only include variables that we think will truly be meaningful to safeguard against this issue.  There are some other strategies, one of which utilizes a technique called PCA which we will discuss later.
