---
title: "Camara_Vogt_Project"
author: "Johnny Vogt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Library and ENV Variables

##Libraries
```{r Libraries, include=FALSE}
library(tidyverse)
library(caret)
library(ggplot2)
library(skimr)
library(corrplot)
library(naniar)
library(GGally)
library(readr)
library(dplyr)
library(glmnet)
library(scales)
```

##Importing the Data
```{r Data Import, include=FALSE}
setwd(dir =  "/Users/Owner/git/School/DS6372_Applied_Statistics/Project1/")
auto <- read.csv(file = "AutoMSRP_Reduced.csv")
head(auto)
str(auto)
glimpse(auto)
```

##Cleaning the Data
Number of Door (impute)
Engine HP (impute)
Engine Cylinder (impute)
Market Category (Has N/A, but keeping them as a category )

Will need to create another model without outliers
```{r Cleaning Data, include=FALSE}
colSums(is.na(auto))
vis_miss(auto)

summary(auto$Number.of.Doors)  # probably Use Median
summary(auto$Engine.HP) # Can Use Mean
summary(auto$Engine.Cylinders) #Probably Median

auto_omit <-na.omit(auto) #Omitting the missing Values ( will look into using imputation)
vis_miss(auto_omit)

auto_omit_continous <- auto_omit[, c("MSRP", "Year", "Engine.HP", "city.mpg")]

auto_omit <- auto_omit %>%  #Factoring
  mutate(
    Make = as.factor(Make),
    `Engine.Fuel.Type` = as.factor(`Engine.Fuel.Type`),
    `Transmission.Type` = as.factor(`Transmission.Type`),
    `Market.Category` = as.factor(`Market.Category`),
    `Vehicle.Size` = as.factor(`Vehicle.Size`),
    `Vehicle.Style` = as.factor(`Vehicle.Style`)
#   `Engine.Cylinders` = as.factor(`Engine.Cylinders`),
#  `Year` = as.factor(`Year`),
#   `Driven_Wheels` = as.factor(`Driven_Wheels`)
  )

##Removing cols: Highway MPg, Cylinders, Vehicle Size, Doors, and make
# List of columns to remove
drop_cols <- c("highway.MPG", "Engine.Cylinders", "Vehicle.Size", "Number.of.Doors", "Make")

# Remove columns using subset
auto_omit_reduced <- auto_omit[, !names(auto_omit) %in% drop_cols]

#Removing Outliers for better modeling
cutoff <- quantile(auto_omit_reduced$MSRP, 0.99)
auto_omit_reduced_no_outliers <- auto_omit_reduced[auto_omit_reduced$MSRP <= cutoff,]

#Identify categorical columns (factor or character)
cat_cols <- names(auto_omit_reduced)[sapply(auto_omit_reduced, function(x) is.factor(x) || is.character(x))]

auto_omit <- auto_omit %>%
  mutate(MPG_bin = cut(city.mpg, breaks = c(0, 15, 20, 25, 30, 35, 100),
                       labels = c("<15","15-20","20-25","25-30","30-35","35+")))

auto_omit_reduced_no_outliers <- auto_omit_reduced_no_outliers %>%
  mutate(MPG_bin = cut(city.mpg, breaks = c(0, 15, 20, 25, 30, 35, 100),
                       labels = c("<15","15-20","20-25","25-30","30-35","35+")))
```

#EDA
##Data Summary and Early Relationship
```{r Early EDA}
#Stats on Missing data
na_rows <- auto %>% filter(if_any(everything(), is.na))
n_removed <- nrow(na_rows)
cat("Rows removed due to NA:", n_removed, "\n")
summary(na_rows)
skim(na_rows)

skimr::skim(auto_omit)

ggpairs(auto_omit_continous)

#BoxPlots
#ggplot(auto_omit, aes(x= Make, y= MSRP)) + 
#  geom_boxplot(outlier.shape = NA) + 
#  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
#  labs(title = "Boxplot of MSRP by Car Make", x = "Make", y = "MSRP (USD)") +
#  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(auto_omit, aes(x= Engine.Fuel.Type, y= MSRP)) +  ### Empty??
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "Boxplot of MSRP by Engine Fuel Type", x = "Engine Fuel Type", y = "MSRP (USD)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(auto_omit, aes(x= Transmission.Type, y= MSRP)) + #Unknown LOL 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "Boxplot of MSRP by Transmission", x = "Transmission Type", y = "MSRP (USD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(auto_omit, aes(x= Transmission.Type, y= MSRP)) + #Unknown LOL 
  geom_boxplot(outlier.shape = NA) + 
  labs(title = "Boxplot of MSRP by Transmission", x = "Transmission Type", y = "MSRP (USD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(auto_omit, aes(x= Market.Category, y= MSRP)) +  
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "Boxplot of MSRP by Market Category", x = "Market Category", y = "MSRP (USD)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#ggplot(auto_omit, aes(x= Vehicle.Size, y= MSRP)) + 
#  geom_boxplot(outlier.shape = NA) + 
#  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
#  labs(title = "Boxplot of MSRP by Vehicle Size", x = "Vehicle Size", y = "MSRP (USD)") +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(auto_omit, aes(x= Vehicle.Style, y= MSRP)) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "Boxplot of MSRP by Vehicle Style", x = "Vehicle Style", y = "MSRP (USD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Wayyyy too cluttered
ggplot(auto_omit, aes(x = Vehicle.Style, y = MSRP, fill = Engine.Fuel.Type)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "MSRP by Vehicle Style (colored by Fuel Type)", x = "Vehicle Style", y = "MSRP (USD)", fill = "Fuel Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(auto_omit, aes(x = Vehicle.Style, y = MSRP, fill = MPG_bin)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
  labs(title = "MSRP by Vehicle Style (colored by city MPG group)", x = "Vehicle Style", y = "MSRP (USD)", fill = "City MPG") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggplot(auto_omit, aes(x= Engine.Cylinders, y= MSRP)) + 
#  geom_boxplot(outlier.shape = NA) + 
#  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
#  labs(title = "Boxplot of MSRP by Engine Cylinder", x = "Engine Cylinder", y = "MSRP (USD)") +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggplot(auto_omit, aes(x= Year, y= MSRP)) + 
#  geom_boxplot(outlier.shape = NA) + 
#  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
#  labs(title = "Boxplot of MSRP by Year", x = "Year", y = "MSRP (USD)") +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggplot(auto_omit, aes(x= Driven_Wheels, y= MSRP)) + 
#  geom_boxplot(outlier.shape = NA) + 
#  coord_cartesian(ylim = quantile(auto_omit$MSRP, c(0.05, 0.95))) +
#  labs(title = "Boxplot of MSRP by Drive Type", x = "Drive Type", y = "MSRP (USD)") +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ScatterPlots
ggplot(auto_omit, aes(x = `Engine.HP`, y = MSRP)) +
  geom_point() +
  geom_smooth(method='lm')

ggplot(auto_omit_reduced_no_outliers, aes(x = `Engine.HP`, y = MSRP, color = MPG_bin)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = 'lm', se = FALSE, color = "black") +
  labs(
    title = "Engine HP vs MSRP Colored by City MPG Bin",
    x = "Engine HP",
    y = "MSRP (USD)",
    color = "City MPG Bin"
  ) +
  scale_y_continuous(labels = scales::comma)+
  theme_minimal()

ggplot(auto_omit, aes(x = `highway.MPG`, y = MSRP)) +
  geom_point() +
  geom_smooth(method='lm')

ggplot(auto_omit_reduced_no_outliers, aes(x = city.mpg, y = MSRP)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_smooth(method = 'lm', se = TRUE, color = "darkred", fill = "pink") +
  labs(
    title = "Relationship Between City MPG and Car Price (MSRP)",
    subtitle = "Most cars with high city MPG have lower MSRPs; exceptions may be electric or luxury models",
    x = "City MPG (Miles per Gallon)",
    y = "MSRP (USD)",
    caption = "Source: AutoMSRP_Reduced.csv"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()

ggplot(auto_omit, aes(x = `Year`, y = MSRP)) +
  geom_point() +
  geom_smooth(method='lm')

ggplot(auto_omit, aes(x = `Engine.Cylinders`, y = MSRP)) +
  geom_point() +
  geom_smooth(method='lm')

#Corr Matrix
cor_matrix <- cor(auto_omit_continous, use = "complete.obs")
print(cor_matrix)
corrplot(cor_matrix, method = "color")

#Categorical Var. Fun Stats
category_summary <- function(df, var) {
  tab <- table(df[[var]])
  data.frame(
    Variable      = var,
    N_Categories  = length(tab),
    Most_Common   = names(tab)[which.max(tab)],
    N_Most_Common = max(tab),
    Least_Common  = names(tab)[which.min(tab)],
    N_Least_Common = min(tab),
    stringsAsFactors = FALSE
  )
}
cat_summary <- do.call(rbind, lapply(cat_cols, function(x) category_summary(auto_omit_reduced_no_outliers, x)))

print(cat_summary)


#Removing MPG_bin for modeling purposes
auto_omit <- dplyr::select(auto_omit, -MPG_bin)
auto_omit_reduced_no_outliers <- dplyr::select(auto_omit_reduced_no_outliers, -MPG_bin)
```

#Feature Selection and Ojective 1 Model
##Using Domain Knowledge to remove Attributes
- City and Highway MPG have high correlation and measure very similar things
- Engine HP and Cylinder are closely related
- Vehicle Style and Size are also 
- No. of Doors could also be connected to Style
- Market Category and Make tend to be closely related as well. 

Removing Highway.MPG, Engine.Cylinder, Vehicle.Size, No. of Doors, Make (The new DF will be created up above)


```{r}
# Full Model (all predictors) 
full.model <- lm(MSRP ~ ., data = auto_omit_reduced)
full.model_noout <- lm(MSRP ~ ., data = auto_omit_reduced_no_outliers)

# Null Model (only intercept)
null.model <- lm(MSRP ~ 1, data = auto_omit_reduced)
null.model_noout <- lm(MSRP ~ 1, data = auto_omit_reduced_no_outliers)

# Forward Selection 
#MSRP ~ Market.Category + Engine.HP + Driven_Wheels + Transmission.Type + Vehicle.Style + city.mpg + Engine.Fuel.Type + Year #AIC 247273.4 (Noout)
forward <- step(null.model, scope = formula(full.model), direction = "forward")
forward_noout <- step(null.model, scope = formula(full.model_noout), direction = "forward")

# Backward Elimination
#MSRP ~ Market.Category + Engine.HP + Driven_Wheels + Transmission.Type + Vehicle.Style + city.mpg + Engine.Fuel.Type + Year
backward <- step(full.model, direction = "backward")
backward_noout <- step(full.model_noout, direction = "backward")

# Stepwise Selection
#MSRP ~ Market.Category + Engine.HP + Driven_Wheels + Transmission.Type + Vehicle.Style + city.mpg + Engine.Fuel.Type + Year
stepwise <- step(null.model, scope = formula(full.model), direction = "both")
stepwise_noout <- step(null.model, scope = formula(full.model_noout), direction = "both")

#Each of the above feature selection had the exact same Model

#Full Model and full model_noout Assumption
par(mfrow = c(2, 2))
plot(full.model)
plot(full.model_noout)
par(mfrow = c(1, 1))

#penafull.model#penalized selection
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1) 

glmnet.fit<-train(MSRP~.,
               data=auto_omit_reduced,
               method="glmnet",
               trControl=fitControl
               )
opt.pen<-glmnet.fit$finalModel$lambdaOpt 
#coef(glmnet.fit$finalModel,opt.pen)
glmnet.fit

#coefs <- coef(glmnet.fit$finalModel, glmnet.fit$finalModel$lambdaOpt)
#active_predictors <- rownames(coefs)[as.numeric(coefs) != 0]
#active_predictors
plot(glmnet.fit)


glmnet_noout.fit<-train(MSRP~.,
               data=auto_omit_reduced_no_outliers,
               method="glmnet",
               trControl=fitControl
               )
opt.pen<-glmnet_noout.fit$finalModel$lambdaOpt 
#coef(glmnet_noout.fit$finalModel,opt.pen)
glmnet_noout.fit

#coefs_noout <- coef(glmnet_noout.fit$finalModel, glmnet_noout.fit$finalModel$lambdaOpt)
#active_predictors_noout <- rownames(coefs_noout)[as.numeric(coefs_noout) != 0]
#active_predictors_noout

#GLMNET_NOOUT ASSUMPTION
##QQ Plot Normal Distribution is violated, the plot this would also apply for the glmnet WITH the outliers
plot(glmnet_noout.fit, which=5)
preds_noout <- predict(glmnet_noout.fit, newdata = auto_omit_reduced_no_outliers)
residuals_noout <- auto_omit_reduced_no_outliers$MSRP - preds_noout
df_resid_noout <- data.frame(Fitted = as.numeric(preds_noout), Residuals = as.numeric(residuals_noout))
ggplot(df_resid_noout, aes(x=Fitted, y=Residuals)) +
  geom_point() +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Residuals vs Fitted | glmnet")

ggplot(data.frame(resid = resid(glmnet_noout.fit)), aes(sample = resid)) +
  stat_qq() +
  stat_qq_line() +
  labs(title="QQ Plot of Residuals")
```
##Log Transforming MSRP
We will transform the MSRP to resolve some assumption violations. The outliers are so strong that, we've decided to see what it looks like when you transfrom the log transform the msrp
```{r}
# Full Model (all predictors) -- LOG TRANSFORMATION
full.model_log <- lm(log(MSRP) ~ ., data = auto_omit_reduced)
full.model_noout_log <- lm(log(MSRP) ~ ., data = auto_omit_reduced_no_outliers)

# Null Model (only intercept) -- LOG TRANSFORMATION
null.model_log <- lm(log(MSRP) ~ 1, data = auto_omit_reduced)
null.model_noout_log <- lm(log(MSRP) ~ 1, data = auto_omit_reduced_no_outliers)

# Forward Selection -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Market.Category + Engine.HP + Engine.Fuel.Type + Vehicle.Style + city.mpg + Transmission.Type + Driven_Wheels #-20175.37 (NoOut) -20196.35(FULL)
forward_log <- step(null.model_log, scope = formula(full.model_log), direction = "forward")
forward_noout_log <- step(null.model_noout_log, scope = formula(full.model_noout_log), direction = "forward")


# Backward Elimination -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Engine.Fuel.Type + Engine.HP + Transmission.Type + Driven_Wheels + Market.Category + Vehicle.Style + city.mpg #AIC: -20175.37 (NoOut) -20196.35 (FULL)
backward_log <- step(full.model_log, direction = "backward")
backward_noout_log <- step(full.model_noout_log, direction = "backward")


# Stepwise Selection -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Market.Category + Engine.HP + Engine.Fuel.Type + city.mpg + Vehicle.Style + Transmission.Type + Driven_Wheels #AIC -20175.37(no out) -20196.35 (FULL)
stepwise_log <- step(null.model_log, scope = formula(full.model_log), direction = "both")
stepwise_noout_log <- step(null.model_noout_log, scope = formula(full.model_noout_log), direction = "both")


# Assumption Plots for Log-Transformed Models
par(mfrow = c(2, 2))
plot(full.model_log)
plot(full.model_noout_log)
par(mfrow = c(1, 1))

#
# Penalized selection using log(MSRP)
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats = 1)

glmnet_log.fit <- train(log(MSRP) ~ .,
                       data = auto_omit_reduced,
                       method = "glmnet",
                       trControl = fitControl
                      )
opt.pen_log <- glmnet_log.fit$finalModel$lambdaOpt 
coef(glmnet_log.fit$finalModel, opt.pen_log)
glmnet_log.fit
coefs_log <- coef(glmnet_log.fit$finalModel, glmnet_log.fit$finalModel$lambdaOpt)
active_predictors_log <- rownames(coefs_log)[as.numeric(coefs_log) != 0]
active_predictors_log

plot(glmnet_log.fit)

glmnet_noout_log.fit <- train(log(MSRP) ~ .,
                             data = auto_omit_reduced_no_outliers,
                             method = "glmnet",
                             trControl = fitControl
                            )
opt.pen_noout_log <- glmnet_noout_log.fit$finalModel$lambdaOpt 
#coef(glmnet_noout_log.fit$finalModel, opt.pen_noout_log)
glmnet_noout_log.fit
coefs_noout_log <- coef(glmnet_noout_log.fit$finalModel, glmnet_noout_log.fit$finalModel$lambdaOpt)
active_predictors_noout_log <- rownames(coefs_noout_log)[as.numeric(coefs_noout_log) != 0]
#active_predictors_noout_log

# Predict and get residuals for log(MSRP)
preds_noout_log <- predict(glmnet_noout_log.fit, newdata = auto_omit_reduced_no_outliers)
residuals_noout_log <- log(auto_omit_reduced_no_outliers$MSRP) - preds_noout_log
df_resid_noout_log <- data.frame(Fitted = as.numeric(preds_noout_log), Residuals = as.numeric(residuals_noout_log))

# Residuals vs Fitted for log model
library(ggplot2)
ggplot(df_resid_noout_log, aes(x=Fitted, y=Residuals)) +
  geom_point() +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title = "Residuals vs Fitted | glmnet log(MSRP)")

# QQ Plot for log residuals
ggplot(data.frame(resid = residuals_noout_log), aes(sample = resid)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "QQ Plot of Residuals (log MSRP)")



##Summary of coefficients 
summary(full.model_noout_log)
coefs_noout_log
```

#Objective 2 Complex MLR and Non-Parametric
##EDA for Non-Parametric model
#Adding interactions Engine.hp*Vehicle.Style, Engine.HP*Transmission.Type, city.mpg*Engine.Fuel.Type, Engine.hp^2, and city.mpg^2
```{r}
ggplot(auto_omit_reduced_no_outliers, aes(x=Engine.HP, y=log(MSRP))) +
  geom_point() +
  geom_smooth(method='lm')


#Evidence of Interaction HP and Vehicle Style
ggplot(auto_omit_reduced_no_outliers, aes(x = Engine.HP, y = MSRP, color = Vehicle.Style)) +
  geom_point(alpha=0.3) +
  geom_smooth(method = "lm") +
  labs(title = "HP vs MSRP by Vehicle Style")

m1 <- lm(MSRP ~ Engine.HP + Vehicle.Style, data = auto_omit_reduced_no_outliers)
m2 <- lm(MSRP ~ Engine.HP * Vehicle.Style, data = auto_omit_reduced_no_outliers)  # add interaction
anova(m1, m2) #Evidence that the interaction is significant

#Evidence of Interaction HP and Transmission Type
ggplot(auto_omit_reduced_no_outliers, aes(x = Engine.HP, y = MSRP, color = Transmission.Type)) +
  geom_point(alpha=0.3) +
  geom_smooth(method = "lm") +
  labs(title = "HP vs MSRP by Transmission")

m1 <- lm(MSRP ~ Engine.HP + Transmission.Type, data = auto_omit_reduced_no_outliers)
m2 <- lm(MSRP ~ Engine.HP * Transmission.Type, data = auto_omit_reduced_no_outliers)  # add interaction
anova(m1, m2) #Evidence that the interaction is significant

#Fuel Type defnitely has an effect
ggplot(auto_omit_reduced_no_outliers, aes(x = city.mpg, y = MSRP, color = Engine.Fuel.Type)) +
  geom_point(alpha=0.3) +
  geom_smooth(method = "lm") +
  labs(title = "HP vs MSRP by Vehicle Style")

data_no_electric <- filter(auto_omit_reduced_no_outliers, Engine.Fuel.Type != "electric")

ggplot(data_no_electric, aes(x = city.mpg, y = MSRP, color = Engine.Fuel.Type)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  labs(title = "City MPG vs MSRP by Fuel Type")

#Polynomial for HP?
ggplot(auto_omit_reduced_no_outliers, aes(x=Engine.HP, y=log(MSRP))) + 
  geom_point() + 
  geom_smooth(method='loess')

mod_linear <- lm(log(MSRP) ~ Engine.HP, data = auto_omit_reduced_no_outliers)
mod_quad   <- lm(log(MSRP) ~ Engine.HP + I(Engine.HP^2), data = auto_omit_reduced_no_outliers)

# Comparing models AIC is lower for quad, showing that its
AIC(mod_linear, mod_quad)

#PolyNomial for mpg?
ggplot(auto_omit_reduced_no_outliers, aes(x=city.mpg, y=log(MSRP))) + 
  geom_point() + 
  geom_smooth(method='loess')

##Checkwithout electric
ggplot(data_no_electric, aes(x=Engine.HP, y=log(MSRP))) + 
  geom_point() + 
  geom_smooth(method='loess')

mod_linear <- lm(log(MSRP) ~ city.mpg, data = auto_omit_reduced_no_outliers)
mod_quad   <- lm(log(MSRP) ~ city.mpg + I(city.mpg^2), data = auto_omit_reduced_no_outliers)

# Comparing models AIC is lower for quad, showing that its significant to keep it in the model
AIC(mod_linear, mod_quad)
```

#Complex Model
```{r}
# Full model with main effects, polynomial, and interactions
complex_model <- lm(
  log(MSRP) ~ 
    Year +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = auto_omit_reduced
)

complex_model_noout <- lm(
  log(MSRP) ~ 
    Year +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = auto_omit_reduced_no_outliers
)

AIC(complex_model)
AIC(complex_model_noout)
summary(complex_model_noout)

#Complex Model Assumptions
par(mfrow = c(2, 2))
plot(complex_model)
par(mfrow = c(1, 1))
par(mfrow = c(2, 2))
plot(complex_model_noout)
par(mfrow = c(1, 1))
```

#Non-Parametric Model
```{r Non-Parametric Models}
#KNN_Full
# 1. Select features
features <- c("Year", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
knn_data <- auto_omit_reduced %>% select(MSRP, all_of(features))

# 2. Train/test split
set.seed(123)
idx <- createDataPartition(knn_data$MSRP, p = 0.8, list = FALSE)
train_set <- knn_data[idx, ]
test_set  <- knn_data[-idx, ]

# 3. Create dummy variables (one-hot encoding for categorical predictors)
dummies   <- dummyVars(MSRP ~ ., data = train_set)
train_enc <- predict(dummies, newdata = train_set)
test_enc  <- predict(dummies, newdata = test_set)
train_enc <- data.frame(MSRP = train_set$MSRP, train_enc)
test_enc  <- data.frame(MSRP = test_set$MSRP, test_enc)

# 4. Center/scale all numeric variables
preproc     <- preProcess(train_enc, method = c("center", "scale"))
train_prep  <- predict(preproc, newdata = train_enc)
test_prep   <- predict(preproc, newdata = test_enc)

# 5. Fit KNN on processed data
set.seed(123)
knn_fit <- train(
  MSRP ~ .,
  data = train_prep,
  method = "knn",
  tuneLength = 10,
  trControl = trainControl(method = "cv", number = 5)
)
print(knn_fit)

knn_preds <- predict(knn_fit, newdata = test_prep)

ggplot(data = NULL, aes(x = test_prep$MSRP, y = knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Actual MSRP",
    y = "Predicted MSRP (KNN)",
    title = "Predicted vs Actual MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = knn_preds, y = test_prep$MSRP - knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(
    x = "Predicted MSRP",
    y = "Residuals (Actual - Predicted)",
    title = "Residuals vs Predicted MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = test_prep$MSRP - knn_preds)) +
  geom_histogram(bins = 40, fill = "#56B4E9", color = "white") +
  labs(
    x = "Prediction Error (Residuals)",
    y = "Count",
    title = "Distribution of Prediction Errors (KNN)"
  )

#KNN_NoOut

# 1. Select features
features <- c("Year", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
knn_data <- auto_omit_reduced_no_outliers %>% select(MSRP, all_of(features))

# 2. Train/test split
set.seed(123)
idx <- createDataPartition(knn_data$MSRP, p = 0.8, list = FALSE)
train_set <- knn_data[idx, ]
test_set  <- knn_data[-idx, ]

# 3. Create dummy variables (one-hot encoding for categorical predictors)
dummies   <- dummyVars(MSRP ~ ., data = train_set)
train_enc <- predict(dummies, newdata = train_set)
test_enc  <- predict(dummies, newdata = test_set)
train_enc <- data.frame(MSRP = train_set$MSRP, train_enc)
test_enc  <- data.frame(MSRP = test_set$MSRP, test_enc)

# 4. Center/scale all numeric variables
preproc     <- preProcess(train_enc, method = c("center", "scale"))
train_prep  <- predict(preproc, newdata = train_enc)
test_prep   <- predict(preproc, newdata = test_enc)

# 5. Fit KNN on processed data
set.seed(123)
knn_fit <- train(
  MSRP ~ .,
  data = train_prep,
  method = "knn",
  tuneLength = 10,
  trControl = trainControl(method = "cv", number = 5)
)
print(knn_fit)

knn_preds <- predict(knn_fit, newdata = test_prep)

ggplot(data = NULL, aes(x = test_prep$MSRP, y = knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Actual MSRP",
    y = "Predicted MSRP (KNN)",
    title = "Predicted vs Actual MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = knn_preds, y = test_prep$MSRP - knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(
    x = "Predicted MSRP",
    y = "Residuals (Actual - Predicted)",
    title = "Residuals vs Predicted MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = test_prep$MSRP - knn_preds)) +
  geom_histogram(bins = 40, fill = "#56B4E9", color = "white") +
  labs(
    x = "Prediction Error (Residuals)",
    y = "Count",
    title = "Distribution of Prediction Errors (KNN)"
  )

#RandomForest_Full

features <- c("Year", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
rf_data <- auto_omit_reduced[, c("MSRP", features)]
rf_data_noout <- auto_omit_reduced_no_outliers[, c("MSRP", features)]

set.seed(123)
idx <- createDataPartition(rf_data$MSRP, p = 0.8, list = FALSE)
train_set <- rf_data[idx, ]
test_set <- rf_data[-idx, ]

set.seed(123)
rf_fit <- train(
  MSRP ~ .,
  data = train_set,
  method = "rf",
  trControl = trainControl(method = "cv", number = 3),  # 5-fold CV for tuning
  ntree = 100  # Optional: set number of trees
)
print(rf_fit)


set.seed(123)
idx <- createDataPartition(rf_data_noout$MSRP, p = 0.8, list = FALSE)
rf_noouttrain_set <- rf_data_noout[idx, ]
rf_noouttest_set <- rf_data_noout[-idx, ]

set.seed(123)
rf_fit_noout <- train(
  MSRP ~ .,
  data = rf_noouttrain_set,
  method = "rf",
  trControl = trainControl(method = "cv", number = 3),  # 5-fold CV for tuning
  ntree = 100  # Optional: set number of trees
)
print(rf_fit_noout)

#Non-Parametric Summary
summary(knn_fit)
summary(rf_fit_noout$coefs)
```


Running a prediction Model that splits the data into train and test set
```{r Prediction Function Simple MLR}

set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced[train_idx, ]
test_set  <- auto_omit_reduced[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced)[sapply(auto_omit_reduced, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (all variables except MSRP) - predicting log(MSRP)
predictors <- setdiff(names(auto_omit_reduced), "MSRP")
formula <- as.formula(paste("log(MSRP) ~", paste(predictors, collapse = "+")))
lm_model <- lm(formula, data = train_set)

#Handle log scale properly
log_predictions <- predict(lm_model, newdata = test_set)  # These are in LOG scale

# Create both versions for proper evaluation
test_set$Pred_log_MSRP <- log_predictions                    # Keep log scale predictions
test_set$Pred_MSRP <- exp(log_predictions)                  # Back-transform to dollar scale

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

#Calculate on both scales for proper interpretation

# === APPROACH 1: Original Dollar Scale (Business Interpretation) ===
rmse_dollars <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae_dollars  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2_dollars   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

# === APPROACH 2: Log Scale (Statistically Appropriate) ===
actual_log <- log(test_set$MSRP[valid])
rmse_log <- sqrt(mean((actual_log - test_set$Pred_log_MSRP[valid])^2))
mae_log  <- mean(abs(actual_log - test_set$Pred_log_MSRP[valid]))
r2_log   <- cor(actual_log, test_set$Pred_log_MSRP[valid])^2

# Convert log RMSE to percentage error for easy interpretation
percentage_error <- (exp(rmse_log) - 1) * 100

cat("\n==== Results for Full Data ====\n")
cat("=====================================================\n")
cat("=== DOLLAR SCALE EVALUATION (Business Perspective) ===\n")
cat("Test RMSE (dollars): $", format(round(rmse_dollars), big.mark = ","), "\n")
cat("Test MAE (dollars):  $", format(round(mae_dollars), big.mark = ","), "\n")
cat("Test R-squared:      ", round(r2_dollars, 4), "\n")
cat("\n=== LOG SCALE EVALUATION (Statistical Perspective) ===\n")
cat("Test RMSE (log):     ", round(rmse_log, 4), "\n")
cat("Test MAE (log):      ", round(mae_log, 4), "\n")
cat("Test R-squared:      ", round(r2_log, 4), "\n")
cat("Typical % Error:     ±", round(percentage_error, 1), "%\n")
cat("=====================================================\n")

#Multiple properly scaled visualizations

# Plot 1: Original Dollar Scale (Actual $ vs Predicted $)
p1 <- ggplot(test_set[valid, ], aes(x = MSRP, y = Pred_MSRP)) +
  geom_point(alpha = 0.6, color = "steelblue", size = 2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Dollar Scale",
       subtitle = paste0("RMSE: $", format(round(rmse_dollars), big.mark = ","), 
                        " | R² = ", round(r2_dollars, 3)),
       x = "Actual MSRP ($)",
       y = "Predicted MSRP ($)") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Plot 2: Log Scale (Both axes in log scale - STATISTICALLY CORRECT)
p2 <- ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_log_MSRP)) +
  geom_point(alpha = 0.6, color = "darkgreen", size = 2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Log Scale",
       subtitle = paste0("RMSE: ", round(rmse_log, 3), 
                        " | Typical Error: ±", round(percentage_error, 1), "%"),
       x = "Actual log(MSRP)",
       y = "Predicted log(MSRP)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Plot 3: Residuals vs Fitted (Proper diagnostic plot for log model)
residuals_log <- log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]
fitted_log <- test_set$Pred_log_MSRP[valid]

p3 <- ggplot(data.frame(fitted = fitted_log, residuals = residuals_log), 
             aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6, color = "purple", size = 2) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  geom_smooth(se = FALSE, color = "blue", method = "loess", size = 1) +
  labs(title = "Residuals vs Fitted Values: Log Scale Model",
       subtitle = "Diagnostic plot for checking model assumptions",
       x = "Fitted log(MSRP)",
       y = "Residuals") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Display all plots
print(p1)
print(p2)
print(p3)

# EXPLANATION OF CORRECTIONS
cat("\n==== KEY CORRECTIONS MADE ====\n")
cat("1. PREDICTION: exp(log_predictions) to convert back to dollars\n")
cat("2. METRICS: Calculated on both scales for proper interpretation\n")
cat("3. PLOTTING: Fixed axis scales and labels to match data\n")
cat("4. DIAGNOSTICS: Added proper residual plot for log model\n")
cat("\n==== INTERPRETATION GUIDE ====\n")
cat("• Use DOLLAR metrics for business decisions\n")
cat("• Use LOG metrics for statistical model comparison\n")
cat("• Percentage error gives intuitive sense of accuracy\n")
cat("• Residual plot shows if model assumptions are met\n")

# BONUS: Model summary for additional insights
cat("\n==== MODEL SUMMARY ====\n")
print(summary(lm_model))


#No Outliers
set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced_no_outliers$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced_no_outliers[train_idx, ]
test_set  <- auto_omit_reduced_no_outliers[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced_no_outliers)[sapply(auto_omit_reduced_no_outliers, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (all variables except MSRP) - predicting log(MSRP)
predictors <- setdiff(names(auto_omit_reduced_no_outliers), "MSRP")
formula <- as.formula(paste("log(MSRP) ~", paste(predictors, collapse = "+")))
lm_model <- lm(formula, data = train_set)

#Handle log scale properly
log_predictions <- predict(lm_model, newdata = test_set)  # These are in LOG scale

# Create both versions for proper evaluation
test_set$Pred_log_MSRP <- log_predictions                    # Keep log scale predictions
test_set$Pred_MSRP <- exp(log_predictions)                  # Back-transform to dollar scale

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

#Calculate on both scales for proper interpretation

# === APPROACH 1: Original Dollar Scale (Business Interpretation) ===
rmse_dollars <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae_dollars  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2_dollars   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

# === APPROACH 2: Log Scale (Statistically Appropriate) ===
actual_log <- log(test_set$MSRP[valid])
rmse_log <- sqrt(mean((actual_log - test_set$Pred_log_MSRP[valid])^2))
mae_log  <- mean(abs(actual_log - test_set$Pred_log_MSRP[valid]))
r2_log   <- cor(actual_log, test_set$Pred_log_MSRP[valid])^2

# Convert log RMSE to percentage error for easy interpretation
percentage_error <- (exp(rmse_log) - 1) * 100

cat("\n==== Results for No Outlier Data ====\n")
cat("=====================================================\n")
cat("=== DOLLAR SCALE EVALUATION (Business Perspective) ===\n")
cat("Test RMSE (dollars): $", format(round(rmse_dollars), big.mark = ","), "\n")
cat("Test MAE (dollars):  $", format(round(mae_dollars), big.mark = ","), "\n")
cat("Test R-squared:      ", round(r2_dollars, 4), "\n")
cat("\n=== LOG SCALE EVALUATION (Statistical Perspective) ===\n")
cat("Test RMSE (log):     ", round(rmse_log, 4), "\n")
cat("Test MAE (log):      ", round(mae_log, 4), "\n")
cat("Test R-squared:      ", round(r2_log, 4), "\n")
cat("Typical % Error:     ±", round(percentage_error, 1), "%\n")
cat("=====================================================\n")

#Multiple properly scaled visualizations

# Plot 1: Original Dollar Scale (Actual $ vs Predicted $)
p1 <- ggplot(test_set[valid, ], aes(x = MSRP, y = Pred_MSRP)) +
  geom_point(alpha = 0.6, color = "steelblue", size = 2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Dollar Scale",
       subtitle = paste0("RMSE: $", format(round(rmse_dollars), big.mark = ","), 
                        " | R² = ", round(r2_dollars, 3)),
       x = "Actual MSRP ($)",
       y = "Predicted MSRP ($)") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Plot 2: Log Scale (Both axes in log scale - STATISTICALLY CORRECT)
p2 <- ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_log_MSRP)) +
  geom_point(alpha = 0.6, color = "darkgreen", size = 2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Log Scale",
       subtitle = paste0("RMSE: ", round(rmse_log, 3), 
                        " | Typical Error: ±", round(percentage_error, 1), "%"),
       x = "Actual log(MSRP)",
       y = "Predicted log(MSRP)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Plot 3: Residuals vs Fitted (Proper diagnostic plot for log model)
residuals_log <- log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]
fitted_log <- test_set$Pred_log_MSRP[valid]

p3 <- ggplot(data.frame(fitted = fitted_log, residuals = residuals_log), 
             aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6, color = "purple", size = 2) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  geom_smooth(se = FALSE, color = "blue", method = "loess", size = 1) +
  labs(title = "Residuals vs Fitted Values: Log Scale Model",
       subtitle = "Diagnostic plot for checking model assumptions",
       x = "Fitted log(MSRP)",
       y = "Residuals") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

# Display all plots
print(p1)
print(p2)
print(p3)

# EXPLANATION OF CORRECTIONS
cat("\n==== KEY CORRECTIONS MADE ====\n")
cat("1. PREDICTION: exp(log_predictions) to convert back to dollars\n")
cat("2. METRICS: Calculated on both scales for proper interpretation\n")
cat("3. PLOTTING: Fixed axis scales and labels to match data\n")
cat("4. DIAGNOSTICS: Added proper residual plot for log model\n")
cat("\n==== INTERPRETATION GUIDE ====\n")
cat("• Use DOLLAR metrics for business decisions\n")
cat("• Use LOG metrics for statistical model comparison\n")
cat("• Percentage error gives intuitive sense of accuracy\n")
cat("• Residual plot shows if model assumptions are met\n")

# BONUS: Model summary for additional insights
cat("\n==== MODEL SUMMARY ====\n")
print(summary(lm_model))

```

```{r Prediction Complex MLR}
set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced[train_idx, ]
test_set  <- auto_omit_reduced[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced)[sapply(auto_omit_reduced, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (predicting log(MSRP))
lm_model <- lm(
  log(MSRP) ~ 
    Year +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = train_set
)

#Back-transform from log scale to original dollar scale
log_predictions <- predict(lm_model, newdata = test_set)
test_set$Pred_MSRP <- exp(log_predictions)  # Convert log predictions back to dollars

#log-scale predictions for log-scale evaluation
test_set$Pred_log_MSRP <- log_predictions

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

#Two approaches for proper evaluation

# Approach 1: Evaluate on Original Dollar Scale (back-transformed predictions)
rmse_dollars <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae_dollars  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2_dollars   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

# Approach 2: Evaluate on Log Scale (both actual and predicted in log scale)
rmse_log <- sqrt(mean((log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid])^2))
mae_log  <- mean(abs(log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]))
r2_log   <- cor(log(test_set$MSRP[valid]), test_set$Pred_log_MSRP[valid])^2

# Convert log-scale RMSE to percentage error for interpretation
percentage_error <- (exp(rmse_log) - 1) * 100

cat("\n==== Results for Full Data ====\n")
cat("=== Original Dollar Scale Evaluation ===\n")
cat("Test RMSE (dollars):", round(rmse_dollars, 2), "\n")
cat("Test MAE (dollars):", round(mae_dollars, 2), "\n")
cat("Test R-squared:", round(r2_dollars, 4), "\n")

cat("\n=== Log Scale Evaluation ===\n")
cat("Test RMSE (log scale):", round(rmse_log, 4), "\n")
cat("Test MAE (log scale):", round(mae_log, 4), "\n")
cat("Test R-squared:", round(r2_log, 4), "\n")
cat("Typical percentage error: ±", round(percentage_error, 1), "%\n")

# Two plots

# Plot 1: Original Dollar Scale (Actual vs Back-transformed Predictions)
p1 <- ggplot(test_set[valid, ], aes(x = MSRP, y = Pred_MSRP)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Original Dollar Scale",
       subtitle = paste("RMSE: $", scales::comma(round(rmse_dollars)), 
                       " | R² = ", round(r2_dollars, 3)),
       x = "Actual MSRP ($)",
       y = "Predicted MSRP ($)") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# Plot 2: Log Scale (Both axes in log scale)
p2 <- ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_log_MSRP)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Log Scale",
       subtitle = paste("RMSE: ", round(rmse_log, 3), 
                       " | Typical Error: ±", round(percentage_error, 1), "%"),
       x = "Actual log(MSRP)",
       y = "Predicted log(MSRP)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# Display both plots
print(p1)
print(p2)

# BONUS: Residual Analysis on Log Scale (proper for this model)
residuals_log <- log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]

p3 <- ggplot(data.frame(fitted = test_set$Pred_log_MSRP[valid], 
                       residuals = residuals_log), 
            aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  geom_smooth(se = FALSE, color = "blue", method = "loess") +
  labs(title = "Residuals vs Fitted: Log Scale Model",
       subtitle = "Check for patterns indicating model assumptions violations",
       x = "Fitted log(MSRP)",
       y = "Residuals (log scale)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

print(p3)

#No Outliers
set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced_no_outliers$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced_no_outliers[train_idx, ]
test_set  <- auto_omit_reduced_no_outliers[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced_no_outliers)[sapply(auto_omit_reduced_no_outliers, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (predicting log(MSRP))
lm_model <- lm(
  log(MSRP) ~ 
    Year +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = train_set
)

#Back-transform from log scale to original dollar scale
log_predictions <- predict(lm_model, newdata = test_set)
test_set$Pred_MSRP <- exp(log_predictions)  # Convert log predictions back to dollars

#log-scale predictions for log-scale evaluation
test_set$Pred_log_MSRP <- log_predictions

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

#Two approaches for proper evaluation

# Approach 1: Evaluate on Original Dollar Scale (back-transformed predictions)
rmse_dollars <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae_dollars  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2_dollars   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

# Approach 2: Evaluate on Log Scale (both actual and predicted in log scale)
rmse_log <- sqrt(mean((log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid])^2))
mae_log  <- mean(abs(log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]))
r2_log   <- cor(log(test_set$MSRP[valid]), test_set$Pred_log_MSRP[valid])^2

# Convert log-scale RMSE to percentage error for interpretation
percentage_error <- (exp(rmse_log) - 1) * 100

cat("\n==== Results for No Outlier Data ====\n")
cat("=== Original Dollar Scale Evaluation ===\n")
cat("Test RMSE (dollars):", round(rmse_dollars, 2), "\n")
cat("Test MAE (dollars):", round(mae_dollars, 2), "\n")
cat("Test R-squared:", round(r2_dollars, 4), "\n")

cat("\n=== Log Scale Evaluation ===\n")
cat("Test RMSE (log scale):", round(rmse_log, 4), "\n")
cat("Test MAE (log scale):", round(mae_log, 4), "\n")
cat("Test R-squared:", round(r2_log, 4), "\n")
cat("Typical percentage error: ±", round(percentage_error, 1), "%\n")

# Two plots

# Plot 1: Original Dollar Scale (Actual vs Back-transformed Predictions)
p1 <- ggplot(test_set[valid, ], aes(x = MSRP, y = Pred_MSRP)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Original Dollar Scale",
       subtitle = paste("RMSE: $", scales::comma(round(rmse_dollars)), 
                       " | R² = ", round(r2_dollars, 3)),
       x = "Actual MSRP ($)",
       y = "Predicted MSRP ($)") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# Plot 2: Log Scale (Both axes in log scale)
p2 <- ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_log_MSRP)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Actual vs Predicted MSRP: Log Scale",
       subtitle = paste("RMSE: ", round(rmse_log, 3), 
                       " | Typical Error: ±", round(percentage_error, 1), "%"),
       x = "Actual log(MSRP)",
       y = "Predicted log(MSRP)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# Display both plots
print(p1)
print(p2)

# BONUS: Residual Analysis on Log Scale (proper for this model)
residuals_log <- log(test_set$MSRP[valid]) - test_set$Pred_log_MSRP[valid]

p3 <- ggplot(data.frame(fitted = test_set$Pred_log_MSRP[valid], 
                       residuals = residuals_log), 
            aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  geom_smooth(se = FALSE, color = "blue", method = "loess") +
  labs(title = "Residuals vs Fitted: Log Scale Model",
       subtitle = "Check for patterns indicating model assumptions violations",
       x = "Fitted log(MSRP)",
       y = "Residuals (log scale)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

print(p3)

cat("\n==== Model Interpretation Guide ====\n")
cat("• Dollar Scale RMSE: How many dollars your predictions are typically off\n")
cat("• Log Scale RMSE: More appropriate for this model's assumptions\n")
cat("• Percentage Error: Easy-to-interpret measure of relative accuracy\n")
cat("• Use log scale metrics for model comparison and diagnostics\n")
cat("• Use dollar scale metrics for practical business interpretation\n")

```

Doing everything with Year as a category in stead of numerical
```{r Year_Group DataSet}
#creating dataset
auto_omit_reduced$Year_Group <- factor(ifelse(auto_omit_reduced$Year < 2001, "Pre-2001", "Post-2001"))
auto_omit_reduced_no_outliers$Year_Group <- factor(ifelse(auto_omit_reduced_no_outliers$Year < 2001, "Pre-2001", "Post-2001"))
auto_omit_reduced <- dplyr::select(auto_omit_reduced, -Year)
auto_omit_reduced_no_outliers <- dplyr::select(auto_omit_reduced_no_outliers, -Year)

#Quick EDA
ggplot(auto_omit_reduced_no_outliers, aes(x = `Year_Group`, y = MSRP)) +
  geom_boxplot() 

ggplot(auto_omit_reduced_no_outliers, aes(x = MSRP, fill = Year_Group)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(title = "MSRP Distribution by Year Group",
       x = "MSRP (USD)",
       fill = "Year Group") +
  theme_minimal()

#Redoing Feature selection:Step Wise, Penalized Regression

##Needing to redo this part cause of the year thingy
# Full Model (all predictors) -- LOG TRANSFORMATION
full.model_log <- lm(log(MSRP) ~ ., data = auto_omit_reduced)
full.model_noout_log <- lm(log(MSRP) ~ ., data = auto_omit_reduced_no_outliers)

# Null Model (only intercept) -- LOG TRANSFORMATION
null.model_log <- lm(log(MSRP) ~ 1, data = auto_omit_reduced)
null.model_noout_log <- lm(log(MSRP) ~ 1, data = auto_omit_reduced_no_outliers)

# Forward Selection -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Market.Category + Engine.HP + Engine.Fuel.Type + Vehicle.Style + city.mpg + Transmission.Type + Driven_Wheels #-20175.37 (NoOut) -20196.35(FULL)
forward_log <- step(null.model_log, scope = formula(full.model_log), direction = "forward")
forward_noout_log <- step(null.model_noout_log, scope = formula(full.model_noout_log), direction = "forward")


# Backward Elimination -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Engine.Fuel.Type + Engine.HP + Transmission.Type + Driven_Wheels + Market.Category + Vehicle.Style + city.mpg #AIC: -20175.37 (NoOut) -20196.35 (FULL)
backward_log <- step(full.model_log, direction = "backward")
backward_noout_log <- step(full.model_noout_log, direction = "backward")


# Stepwise Selection -- LOG TRANSFORMATION
#log(MSRP) ~ Year + Market.Category + Engine.HP + Engine.Fuel.Type + city.mpg + Vehicle.Style + Transmission.Type + Driven_Wheels #AIC -20175.37(no out) -20196.35 (FULL)
stepwise_log <- step(null.model_log, scope = formula(full.model_log), direction = "both")
stepwise_noout_log <- step(null.model_noout_log, scope = formula(full.model_noout_log), direction = "both")


# Assumption Plots for Log-Transformed Models
par(mfrow = c(2, 2))
plot(full.model_log)
plot(full.model_noout_log)
par(mfrow = c(1, 1))

#
# Penalized selection using log(MSRP)
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats = 1)

glmnet_log.fit <- train(log(MSRP) ~ .,
                       data = auto_omit_reduced,
                       method = "glmnet",
                       trControl = fitControl
                      )
opt.pen_log <- glmnet_log.fit$finalModel$lambdaOpt 
#coef(glmnet_log.fit$finalModel, opt.pen_log)
glmnet_log.fit
#coefs_log <- coef(glmnet_log.fit$finalModel, glmnet_log.fit$finalModel$lambdaOpt)
#active_predictors_log <- rownames(coefs_log)[as.numeric(coefs_log) != 0]
#active_predictors_log

plot(glmnet_log.fit)

glmnet_noout_log.fit <- train(log(MSRP) ~ .,
                             data = auto_omit_reduced_no_outliers,
                             method = "glmnet",
                             trControl = fitControl
                            )
opt.pen_noout_log <- glmnet_noout_log.fit$finalModel$lambdaOpt 
#coef(glmnet_noout_log.fit$finalModel, opt.pen_noout_log)
glmnet_noout_log.fit
#coefs_noout_log <- coef(glmnet_noout_log.fit$finalModel, glmnet_noout_log.fit$finalModel$lambdaOpt)
#active_predictors_noout_log <- rownames(coefs_noout_log)[as.numeric(coefs_noout_log) != 0]
#active_predictors_noout_log

# Predict and get residuals for log(MSRP)
preds_noout_log <- predict(glmnet_noout_log.fit, newdata = auto_omit_reduced_no_outliers)
residuals_noout_log <- log(auto_omit_reduced_no_outliers$MSRP) - preds_noout_log
df_resid_noout_log <- data.frame(Fitted = as.numeric(preds_noout_log), Residuals = as.numeric(residuals_noout_log))

# Residuals vs Fitted for log model
library(ggplot2)
ggplot(df_resid_noout_log, aes(x=Fitted, y=Residuals)) +
  geom_point() +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title = "Residuals vs Fitted | glmnet log(MSRP)")

# QQ Plot for log residuals
ggplot(data.frame(resid = residuals_noout_log), aes(sample = resid)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "QQ Plot of Residuals (log MSRP)")

#Non Parametric Stuff
#KNN_Full
# 1. Select features
features <- c("Year_Group", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
knn_data <- auto_omit_reduced %>% select(MSRP, all_of(features))

# 2. Train/test split
set.seed(123)
idx <- createDataPartition(knn_data$MSRP, p = 0.8, list = FALSE)
train_set <- knn_data[idx, ]
test_set  <- knn_data[-idx, ]

# 3. Create dummy variables (one-hot encoding for categorical predictors)
dummies   <- dummyVars(MSRP ~ ., data = train_set)
train_enc <- predict(dummies, newdata = train_set)
test_enc  <- predict(dummies, newdata = test_set)
train_enc <- data.frame(MSRP = train_set$MSRP, train_enc)
test_enc  <- data.frame(MSRP = test_set$MSRP, test_enc)

# 4. Center/scale all numeric variables
preproc     <- preProcess(train_enc, method = c("center", "scale"))
train_prep  <- predict(preproc, newdata = train_enc)
test_prep   <- predict(preproc, newdata = test_enc)

# 5. Fit KNN on processed data
set.seed(123)
knn_fit <- train(
  MSRP ~ .,
  data = train_prep,
  method = "knn",
  tuneLength = 10,
  trControl = trainControl(method = "cv", number = 5)
)
print(knn_fit)

knn_preds <- predict(knn_fit, newdata = test_prep)

ggplot(data = NULL, aes(x = test_prep$MSRP, y = knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Actual MSRP",
    y = "Predicted MSRP (KNN)",
    title = "Predicted vs Actual MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = knn_preds, y = test_prep$MSRP - knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(
    x = "Predicted MSRP",
    y = "Residuals (Actual - Predicted)",
    title = "Residuals vs Predicted MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = test_prep$MSRP - knn_preds)) +
  geom_histogram(bins = 40, fill = "#56B4E9", color = "white") +
  labs(
    x = "Prediction Error (Residuals)",
    y = "Count",
    title = "Distribution of Prediction Errors (KNN)"
  )

#KNN_NoOut

# 1. Select features
features <- c("Year_Group", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
knn_data <- auto_omit_reduced_no_outliers %>% select(MSRP, all_of(features))

# 2. Train/test split
set.seed(123)
idx <- createDataPartition(knn_data$MSRP, p = 0.8, list = FALSE)
train_set <- knn_data[idx, ]
test_set  <- knn_data[-idx, ]

# 3. Create dummy variables (one-hot encoding for categorical predictors)
dummies   <- dummyVars(MSRP ~ ., data = train_set)
train_enc <- predict(dummies, newdata = train_set)
test_enc  <- predict(dummies, newdata = test_set)
train_enc <- data.frame(MSRP = train_set$MSRP, train_enc)
test_enc  <- data.frame(MSRP = test_set$MSRP, test_enc)

# 4. Center/scale all numeric variables
preproc     <- preProcess(train_enc, method = c("center", "scale"))
train_prep  <- predict(preproc, newdata = train_enc)
test_prep   <- predict(preproc, newdata = test_enc)

# 5. Fit KNN on processed data
set.seed(123)
knn_fit <- train(
  MSRP ~ .,
  data = train_prep,
  method = "knn",
  tuneLength = 10,
  trControl = trainControl(method = "cv", number = 5)
)
print(knn_fit)

knn_preds <- predict(knn_fit, newdata = test_prep)

ggplot(data = NULL, aes(x = test_prep$MSRP, y = knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Actual MSRP",
    y = "Predicted MSRP (KNN)",
    title = "Predicted vs Actual MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = knn_preds, y = test_prep$MSRP - knn_preds)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(
    x = "Predicted MSRP",
    y = "Residuals (Actual - Predicted)",
    title = "Residuals vs Predicted MSRP (KNN Regression)"
  ) +
  theme_minimal()

ggplot(data = NULL, aes(x = test_prep$MSRP - knn_preds)) +
  geom_histogram(bins = 40, fill = "#56B4E9", color = "white") +
  labs(
    x = "Prediction Error (Residuals)",
    y = "Count",
    title = "Distribution of Prediction Errors (KNN)"
  )

#RandomForest_Full

features <- c("Year_Group", "Market.Category", "Engine.HP", "Engine.Fuel.Type", 
              "city.mpg", "Vehicle.Style", "Transmission.Type", "Driven_Wheels")
rf_data <- auto_omit_reduced[, c("MSRP", features)]
rf_data_noout <- auto_omit_reduced_no_outliers[, c("MSRP", features)]

set.seed(123)
idx <- createDataPartition(rf_data$MSRP, p = 0.8, list = FALSE)
train_set <- rf_data[idx, ]
test_set <- rf_data[-idx, ]

set.seed(123)
rf_fit <- train(
  MSRP ~ .,
  data = train_set,
  method = "rf",
  trControl = trainControl(method = "cv", number = 3),  # 5-fold CV for tuning
  ntree = 100  # Optional: set number of trees
)
print(rf_fit)


set.seed(123)
idx <- createDataPartition(rf_data_noout$MSRP, p = 0.8, list = FALSE)
rf_noouttrain_set <- rf_data_noout[idx, ]
rf_noouttest_set <- rf_data_noout[-idx, ]

set.seed(123)
rf_fit_noout <- train(
  MSRP ~ .,
  data = rf_noouttrain_set,
  method = "rf",
  trControl = trainControl(method = "cv", number = 3),  # 5-fold CV for tuning
  ntree = 100  # Optional: set number of trees
)
print(rf_fit_noout)

##Prediction on Train/Test Set #####
set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced[train_idx, ]
test_set  <- auto_omit_reduced[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced)[sapply(auto_omit_reduced, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (all variables except MSRP)
#predictors <- setdiff(names(auto_omit_reduced), "MSRP")
#formula <- as.formula(paste("log(MSRP) ~", paste(predictors, collapse = "+")))
lm_model <- lm(
  log(MSRP) ~ 
    Year_Group +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = train_set
)

# Predict
test_set$Pred_MSRP <- predict(lm_model, newdata = test_set)

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

# Metrics (only valid rows)
rmse <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

cat("\n==== Results for Full Data (Year_Group) ====\n")
cat("Test RMSE:", rmse, "\n")
cat("Test MAE:", mae, "\n")
cat("Test R-squared:", r2, "\n")

# Plot
ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_MSRP)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted MSRP: Full Data",
       x = "Actual MSRP",
       y = "Predicted MSRP") +
  theme_minimal()

#No Outliers
set.seed(123)
train_idx <- createDataPartition(auto_omit_reduced_no_outliers$MSRP, p = 0.8, list = FALSE)
train_set <- auto_omit_reduced_no_outliers[train_idx, ]
test_set  <- auto_omit_reduced_no_outliers[-train_idx, ]

# Force categorical/factor levels in test to match train
cat_vars <- names(auto_omit_reduced_no_outliers)[sapply(auto_omit_reduced_no_outliers, is.factor)]
for (v in cat_vars) {
  train_set[[v]] <- factor(train_set[[v]])
  test_set[[v]]  <- factor(test_set[[v]], levels = levels(train_set[[v]]))
}

# Model formula (all variables except MSRP)
#predictors <- setdiff(names(auto_omit_reduced_no_outliers), "MSRP")
#formula <- as.formula(paste("log(MSRP) ~", paste(predictors, collapse = "+")))
lm_model <- lm(
  log(MSRP) ~ 
    Year_Group +
    Market.Category +
    Engine.Fuel.Type +
    Vehicle.Style +
    Transmission.Type +
    Driven_Wheels +
    city.mpg +
    Engine.HP +
    I(Engine.HP^2) +  # Quadratic (polynomial)
    Engine.HP:Vehicle.Style +
    Engine.HP:Transmission.Type +
    city.mpg:Engine.Fuel.Type,
  data = train_set
)

# Predict
test_set$Pred_MSRP <- predict(lm_model, newdata = test_set)

# Restrict to valid predictions
valid <- !is.na(test_set$Pred_MSRP) & !is.na(test_set$MSRP)

# Metrics (only valid rows)
rmse <- sqrt(mean((test_set$MSRP[valid] - test_set$Pred_MSRP[valid])^2))
mae  <- mean(abs(test_set$MSRP[valid] - test_set$Pred_MSRP[valid]))
r2   <- cor(test_set$MSRP[valid], test_set$Pred_MSRP[valid])^2

cat("\n==== Results for No Outlier Data (Year_Group) ====\n")
cat("Test RMSE:", rmse, "\n")
cat("Test MAE:", mae, "\n")
cat("Test R-squared:", r2, "\n")

# Plot
ggplot(test_set[valid, ], aes(x = log(MSRP), y = Pred_MSRP)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted MSRP: No Outlier Data",
       x = "Actual MSRP",
       y = "Predicted MSRP") +
  theme_minimal()
```

