---
title: "Project 2: Adult Income Classification"
author: "Mike Flores and Johnny Vogt"
date: "2025-11-20"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(caret)
library(MASS)
library(class)
library(randomForest)
library(pROC)
library(broom)
library(knitr)
library(kableExtra)
set.seed(1234)
```

# 1. Introduction

This project applies classification methods to the **Adult Income dataset**, a census-based dataset containing approximately 32,000 observations and 14 predictors, with the binary outcome:

- **">50K"**
- **"<=50K"**

The analysis is divided into two major objectives:

**Objective 1:**  
Create a *clean and interpretable* logistic regression model, chosen using AIC, and interpret at least one continuous and one categorical coefficient.

**Objective 2:**  
Build *competing classification models* including:

- Complex logistic regression (interactions or polynomials)
- LDA
- QDA
- kNN
- Random Forest

Performance is compared using:

- Sensitivity  
- Specificity  
- PPV  
- NPV  
- Prevalence  
- ROC  
- AUC  

Validation data is used for all final comparisons.

# 2. Data Loading & Cleaning

```{r load-clean}
adult <- read_csv("adult_income.csv",
                  na = c("?", " ?"))

adult <- adult %>%
  mutate(
    income = factor(income),
    sex    = factor(sex),
    workclass = factor(workclass),
    education = factor(education),
    marital_status = factor(marital_status),
    occupation = factor(occupation),
    relationship = factor(relationship),
    race = factor(race),
    native_country = factor(native_country)
  ) %>%
  drop_na()

glimpse(adult)
```

# 3. Train/Validation Split

```{r split}
train_idx <- createDataPartition(adult$income, p = 0.7, list = FALSE)
train <- adult[train_idx, ]
valid <- adult[-train_idx, ]

nrow(train); nrow(valid)

train %>%
  count(income) %>%
  mutate(prop = n/sum(n))
```

# 4. Exploratory Data Analysis (EDA)

## 4.1 Distribution of Numeric Predictors

```{r eda-num1}
train %>%
  ggplot(aes(x = age, fill = income)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30)
```

```{r eda-num2}
train %>%
  ggplot(aes(x = hours_per_week, fill = income)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30)
```

## 4.2 Categorical Predictors

```{r eda-cat}
train %>%
  count(education, income) %>%
  group_by(education) %>%
  mutate(prop = n/sum(n)) %>%
  ggplot(aes(x = fct_reorder(education, prop), y = prop, fill = income)) +
  geom_col(position = "fill") +
  coord_flip()
```

# 5. Objective 1: Interpretable Logistic Regression

```{r logit1}
logit1 <- glm(
  income ~ age + education + marital_status + occupation +
           relationship + sex + hours_per_week +
           capital_gain + capital_loss,
  data = train,
  family = binomial
)

summary(logit1)
AIC(logit1)

logit1_or <- tidy(logit1, conf.int = TRUE, exponentiate = TRUE)
kable(logit1_or) %>% kable_styling()
```

## 5.1 Interpretation

*(Write your interpretation here after knitting.)*

# 6. Objective 1: Validation Metrics

```{r metrics-helpers}
metrics_from_cm <- function(cm, positive) {
  tp <- cm[positive, positive]
  fn <- sum(cm[, positive]) - tp
  fp <- sum(cm[positive, ]) - tp
  tn <- sum(cm) - tp - fn - fp
  
  tibble(
    Sensitivity = tp / (tp + fn),
    Specificity = tn / (tn + fp),
    PPV = tp / (tp + fp),
    NPV = tn / (tn + fn),
    Prevalence = (tp + fn) / sum(cm)
  )
}
```

```{r logit1-pred}
valid$prob_logit1 <- predict(logit1, newdata = valid, type = "response")
valid$pred_logit1 <- ifelse(valid$prob_logit1 >= 0.5, ">50K", "<=50K") %>%
  factor(levels = levels(valid$income))

cm1 <- table(Predicted = valid$pred_logit1, Actual = valid$income)
met1 <- metrics_from_cm(cm1, positive = ">50K")

kable(cm1)
met1

roc1 <- roc(valid$income, valid$prob_logit1)
auc(roc1)

plot(roc1, main = "ROC: Logistic Regression (Objective 1)")
```

# 7. Objective 2: Complex Logistic Regression

```{r logit2}
logit2 <- glm(
  income ~ age + I(age^2) +
           education + marital_status +
           occupation + relationship +
           sex * hours_per_week +
           capital_gain + capital_loss,
  data = train,
  family = binomial
)

AIC(logit1, logit2)
```

```{r logit2-pred}
valid$prob_logit2 <- predict(logit2, newdata = valid, type = "response")
valid$pred_logit2 <- ifelse(valid$prob_logit2 >= 0.5, ">50K", "<=50K") %>%
  factor(levels = levels(valid$income))

cm2 <- table(Predicted = valid$pred_logit2, Actual = valid$income)
met2 <- metrics_from_cm(cm2, positive = ">50K")

kable(cm2)
met2

roc2 <- roc(valid$income, valid$prob_logit2)
auc(roc2)
```

# 8. Threshold Selection

```{r threshold}
coords2 <- coords(roc2, x = "best", best.method = "youden",
                  ret = c("threshold","sensitivity","specificity"))
coords2

best_thresh <- coords2["threshold"]

valid$pred_logit2_best <- ifelse(valid$prob_logit2 >= best_thresh,
                                 ">50K", "<=50K") %>%
  factor(levels = levels(valid$income))

cm2b <- table(Predicted = valid$pred_logit2_best, Actual = valid$income)
met2b <- metrics_from_cm(cm2b, positive = ">50K")
met2b
```

# 9. LDA and QDA

```{r lda-qda}
dmy <- dummyVars(income ~ ., data = train)
train_x <- predict(dmy, newdata = train) %>% as.data.frame()
train_y <- train$income

valid_x <- predict(dmy, newdata = valid) %>% as.data.frame()
valid_y <- valid$income

lda_fit <- lda(train_x, grouping = train_y)
lda_prob <- predict(lda_fit, valid_x)$posterior[, ">50K"]
lda_pred <- ifelse(lda_prob >= 0.5, ">50K", "<=50K") %>% factor(levels = levels(valid_y))

cm_lda <- table(Predicted = lda_pred, Actual = valid_y)
met_lda <- metrics_from_cm(cm_lda, positive = ">50K")
roc_lda <- roc(valid_y, lda_prob)
auc(roc_lda)

qda_fit <- qda(train_x, grouping = train_y)
qda_prob <- predict(qda_fit, valid_x)$posterior[, ">50K"]
qda_pred <- ifelse(qda_prob >= 0.5, ">50K", "<=50K") %>% factor(levels = levels(valid_y))

cm_qda <- table(Predicted = qda_pred, Actual = valid_y)
met_qda <- metrics_from_cm(cm_qda, positive = ">50K")
roc_qda <- roc(valid_y, qda_prob)
auc(roc_qda)
```

# 10. kNN Model

```{r knn}
preproc <- preProcess(train_x, method = c("center","scale"))
train_x_sc <- predict(preproc, train_x)
valid_x_sc <- predict(preproc, valid_x)

knn_pred <- knn(train_x_sc, valid_x_sc, cl = train_y, k = 15)

cm_knn <- table(Predicted = knn_pred, Actual = valid_y)
met_knn <- metrics_from_cm(cm_knn, positive = ">50K")
met_knn
```

# 11. Random Forest (Nonparametric)

```{r rf}
set.seed(1234)
rf_fit <- randomForest(
  x = train_x,
  y = train_y,
  ntree = 300,
  mtry = floor(sqrt(ncol(train_x))),
  importance = TRUE
)

rf_prob <- predict(rf_fit, valid_x, type = "prob")[, ">50K"]
rf_pred <- ifelse(rf_prob >= 0.5, ">50K", "<=50K") %>% factor(levels = levels(valid_y))

cm_rf <- table(Predicted = rf_pred, Actual = valid_y)
met_rf <- metrics_from_cm(cm_rf, positive = ">50K")

roc_rf <- roc(valid_y, rf_prob)
auc(roc_rf)
```

# 12. ROC Comparison

```{r roc-compare}
plot(roc1, col="black", main = "ROC Comparison")
plot(roc2, col="blue", add=TRUE)
plot(roc_lda, col="red", add=TRUE)
plot(roc_qda, col="green", add=TRUE)
plot(roc_rf, col="purple", add=TRUE)

legend("bottomright",
       legend = c("Logit1","Logit2","LDA","QDA","RF"),
       col = c("black","blue","red","green","purple"),
       lwd = 2)
```

# 13. Comparison Table

```{r compare-table}
results <- bind_rows(
  tibble(Model = "Logit1", met1),
  tibble(Model = "Logit2", met2),
  tibble(Model = "Logit2_best", met2b),
  tibble(Model = "LDA", met_lda),
  tibble(Model = "QDA", met_qda),
  tibble(Model = "kNN_15", met_knn),
  tibble(Model = "RandomForest", met_rf)
)

kable(results) %>% kable_styling(full_width = FALSE)
```

# 14. Conclusions

Write your conclusions here.

